---
title: "NewFunc"
author: "SA22231079"
date: "2023.12.05"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{NewFunc}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview
My topic revolves around ion beam control and I need to simulate ion trajectories under various field effects, so solving the differential equations for the ion trajectories helps to understand the ion motion. This time, I took the opportunity of learning Rcpp to implement two simple types of numerical methods for differential equations, Euler's method and Lungkuta's method, using R and Rcpp respectively:

The fourth-order Dragon Gekkuta method takes the following form

$$
\begin{cases}y_{n+1}=y_n+\frac{1}{6}\big(k_1+2k_2+2k_3+k_4\big)\\k_1=hf(x_n,y_n)\\k_2=hf(x_n+\frac{h}{2},y_n+\frac{k_1}{2})\\k_3=hf(x_n+\frac{h}{2},y_n+\frac{k_2}{2})\\k_4=hf(x_n+h,y_n+k_3)\end{cases}
$$

The forward Eulerian principle is then a simple single-step integral approximation
$$\begin{aligned}y_{i+1}=y_i+hf(t_i,y_i)&&\end{aligned}$$
$$
y_{i+1}=y_i+hky_i=(1+kh)^{i+1}
$$
where k is the slope

## Rongo Kuta


```{r}
#R：
RK4 <- function(dynamics, y0, t0 = 0, tn = 1, n = 1000)
{
  h <- (tn-t0)/n
  y <- matrix(0, n+1)
  # n+1 points, n sub-intervals
  tt <- seq(t0, tn, length.out = n+1)
  y[1] <- y0
  for(i in 2:(n+1))
  {
    k1 <- dynamics(tt[i-1], y[i-1])
    k2 <- dynamics(tt[i-1]+0.5*h, y[i-1]+0.5*h*k1)
    k3 <- dynamics(tt[i-1]+0.5*h, y[i-1]+0.5*h*k2)
    k4 <- dynamics(tt[i-1]+h, y[i-1]+h*k3)
    
    y[i] <- y[i-1] + (h/6)*(k1+2*k2+2*k3+k4)
  }
  return(data.frame(time = tt, state = y))
}
f1 <- function(t, v) 9.8 - v^2
v1 <- RK4(dynamics = f1, y0 = 0, tn = 1, n = 5000)
plot(v1)
```
C++:
```{Rcpp, cache=TRUE}
#include <Rcpp.h>
using namespace Rcpp;

// Define the system of differential equations
void f(double t, NumericVector y, NumericVector dydt, double g, double c, double m) {
  dydt[0] = y[1];
  dydt[1] = g - (c / m) * y[1] * y[1];
}

// [[Rcpp::export]]
NumericMatrix rungeKutta(double t0, NumericVector y0, double h, int n, double g, double c, double m) {
  int dim = y0.length();
  
  NumericMatrix y(n + 1, dim);
  NumericVector k1(dim), k2(dim), k3(dim), k4(dim);
  
  double t = t0;
  y(0, _) = y0;
  
  for (int i = 0; i < n; i++) {
    f(t, y(i, _), k1, g, c, m);
    f(t + h / 2, y(i, _) + h * k1 / 2, k2, g, c, m);
    f(t + h / 2, y(i, _) + h * k2 / 2, k3, g, c, m);
    f(t + h, y(i, _) + h * k3, k4, g, c, m);
    
    y(i + 1, _) = y(i, _) + h * (k1 + 2 * k2 + 2 * k3 + k4) / 6;
    
    t += h;
  }
  
  return y;
}

```

```{r, rungeKutta()}
library(Rcpp)
ans <- rungeKutta(t0 = 0, y0 = c(0,0), h = 0.0002, n = 10000, g = 9.8, c = 0.5, m = 1.0)
plot(ans)
```
Speed vs:
```{r}
library(microbenchmark)
tm1 <- microbenchmark::microbenchmark(
  rnR = RK4(dynamics = function(t, v) 9.8 - v^2, y0=0, t0 = 0, tn = 1, n = 5000),
  rnC = rungeKutta(t0 = 0, y0 = c(0,0), h = 0.0002, n = 10000, g = 9.8, c = 0.5, m = 1.0)
)
print(summary(tm1)[,c(1,3,5,6)])
```


## Eulerscheme

```{r}
#R：
EulerScheme <- function(dynamics, y0, t0 = 0, tn = 1, n = 1000)
{
  h <- (tn-t0)/n
  y <- matrix(0, n+1)
  # n+1 points, n sub-intervals
  tt <- seq(t0, tn, length.out = n+1)
  y[1] <- y0
  for(i in 2:(n+1))
  {
    y[i] <- y[i-1] + h*dynamics(tt[i-1], y[i-1])
  }
  return(data.frame(time = tt, state = y))
}
f1 <- function(t, v) 9.8 - v^2
v1 <- EulerScheme(dynamics = f1, y0 = 0, tn = 1, n = 5000)
plot(v1)

```
C++:
```{Rcpp, cache = TRUE}
#include <Rcpp.h>
using namespace Rcpp;


// Define the system of differential equations
void f2(double t, NumericVector y, NumericVector dydt, double g, double c, double m) {
  dydt[0] = y[1];
  dydt[1] = g - (c / m) * y[1] * y[1];
}

// [[Rcpp::export]]
NumericMatrix eulerMethod(double t0, NumericVector y0, double h, int n, double g, double c, double m) {
  int dim = y0.length();
  
  NumericMatrix y(n + 1, dim);
  
  double t = t0;
  y(0, _) = y0;
  
  for (int i = 0; i < n; i++) {
    NumericVector dydt(dim);
    f2(t, y(i, _), dydt, g, c, m);
    
    y(i + 1, _) = y(i, _) + h * dydt;
    
    t += h;
  }
  
  return y;
}
```
```{r, eulerMethod()}
library(Rcpp)
ans <- eulerMethod(t0 = 0, y0 = c(0,0), h = 0.0002, n = 10000, g = 9.8, c = 0.5, m = 1.0)
plot(ans)
```

Speed vs:
```{r, EulerScheme(),eulerMethod()}
library(microbenchmark)
tm1 <- microbenchmark::microbenchmark(
  rnR = EulerScheme(dynamics = function(t, v) 9.8 - v^2, y0=0, t0 = 0, tn = 1, n = 5000),
  rnC = eulerMethod(t0 = 0, y0 = c(0,0), h = 0.0002, n = 10000, g = 9.8, c = 0.5, m = 1.0)
)
print(summary(tm1)[,c(1,3,5,6)])
```

